---
title: "Tidyverse Study"
author: "Sejin Lim"
date: '2020 3 18 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Library Packages

```{r}
library(tidyverse)
```


```{r}
write_csv(storms, "data/storms.csv")
```

This is how I import data into my R Notebook. I am going to use the tidyverse pacakage from *readr*, called `read_csv()`

```{r}
storms2 <- read_csv("data/storms.csv")
```

Let's have a look at a storms data

```{r}
storms2
```

```{r}
glimpse(storms2)   # it is like str() prettier version in tidyverse
```

This is how I can look at a vector of my dataframe

```{r}
storms2$status
```

# dplyr

- filter (rows)
- arrange (sort rows by variables)
- select (columns)
- mutate (change cell values)
- Count
- Summarize

```{r}
starwars
```

## Filter

```{r}
#starwars_small <- starwars %>% 
#  filter(!is.na(mass), !is.na(height))
starwars_small <- starwars %>% 
  filter(mass >= 0)
```

## Arrange

```{r}
starwars_small %>% 
  arrange(desc(height), desc(name))
```

## Select

```{r}
starwars_small %>% 
  arrange(species, height) %>% 
  select(name, species, height, mass, birth_year)
```


```{r}
starwars_small %>% 
  select(name, species, height, mass, birth_year) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") 
  
```


## Mutate

```{r}
starwars_small %>% 
  select(name, species, height, mass, birth_year) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") %>% 
  mutate(BMI = round(height / mass, 2))
```


## Count

```{r}
starwars %>% 
  count(mass) %>% 
  tidyr::drop_na() %>% 
  arrange(desc(n))
```


## Summarize

```{r}
starwars %>% 
  filter(!is.na(height)) %>% 
  group_by(species) %>% 
  summarise(Count = n(), mean_ht = mean(height), min_ht = min(height),
            max_ht = max(height), sum_ht = sum(height)) %>% 
  arrange(desc(Count))
```


