---
title: "Tidyverse Study"
author: "Sejin Lim"
date: '2020 3 18 '
output: 
 html_document:
  code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Library Packages

```{r}
library(tidyverse)
```


```{r}
write_csv(storms, "data/storms.csv")
```


This is how I import data into my R Notebook. I am going to use the tidyverse pacakage from *readr*, called `read_csv()`

```{r}
storms2 <- read_csv("data/storms.csv")
```

Let's have a look at a storms data
`str()`, `attributes()`, `glimpse()`

```{r}
glimpse(storms2)   # it is like str() prettier version in tidyverse
```

This is how I can look at a vector of my dataframe

```{r}
head(storms2$status, 10)
```

# dplyr

- tbl_df (convert data frame to one with better printing)
- filter (subset rows)
- arrange (sort rows by variables)
- select (subset columns)
- rename (rename variables)
- mutate (change cell values, add newly created column)
- transmute (create new variable)
- count
- summarize (apply basic function to data)
- conditional excution
- gather (make wide data longer)
- spread (make long data wider)

Read a data 
```{r}
starwars <- tbl_df(starwars)
```

## Filter

```{r}
#starwars_small <- starwars %>% 
#  filter(!is.na(mass), !is.na(height))
starwars_small <- starwars %>% 
  filter(mass >= 0)

starwars_small %>%
  filter(height >=180 & species=="Human" )

```

## Arrange

```{r}
starwars_small %>% 
  arrange(desc(height), desc(name))
```

## Select

`starwars$name` simplifys to vector, but `select(name)` maintains the dataframe shape. 
We can reorder variables using `select()` 

```{r}
starwars_small %>% 
  arrange(species, height) %>% 
  select(name, species, height, mass, birth_year)
```

`height:birth_year` means all the columns next to each other between height and birth_year.
`contains("color")` means all the columns which contains "color"
`starts_with("g")` means all the columns which starts with "g" e.g. gender
`ends_with("color")` means all the columns which ends with "color"

```{r}
starwars_small %>% 
  select(name, species, height:birth_year) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") 

starwars_small %>% 
  select(name, species, height, mass, contains("color")) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") 

starwars_small %>% 
  select(name, species, height, mass, starts_with("g")) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") 
  
```

## Rename

```{r}
starwars_small %>%
  select(starts_with("m")) %>%
  rename("weight" = mass)

```


## Mutate $ Transmute

`mutate`: add newly created column using syntax `mutate(data, newVarName = functionOfData, newVarName2 = functionOfData, ...)`

`transmute`: create new variable : it likes `mutate()` and `select()` together 


```{r}
starwars_small %>% 
  select(name, species, height, mass, birth_year) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") %>% 
  mutate(BMI = round(height / mass, 2))


starwars_small %>% 
  select(name, species, height, mass, birth_year) %>% 
  arrange(species, desc(height), mass, name) %>% 
  filter(species == "Human") %>% 
  transmute(BMI = round(height / mass, 2))
```


## Count

```{r}
starwars %>% 
  count(mass) %>% 
  tidyr::drop_na() %>% 
  arrange(desc(n))
```


## Summarize

`group_by()` : group rows by a variable

```{r}
starwars %>% 
  filter(!is.na(height)) %>% 
  group_by(species) %>% 
  summarise(Count = n(), mean_ht = mean(height), min_ht = min(height),
            max_ht = max(height), sum_ht = sum(height)) %>% 
  arrange(desc(Count))
```

## Conditional Execution

Often want to execute statements conditionally (say to create a new variable)
`if then else` syntax
- `if` condition can only take in a single comparison : `if(condtion){then excute code}`
- `ifelse` is vectorized if statment : `ifelse(vector_condition, if_true_do_this, if_false_do_this)` 

```{r}
mutate(starwars_small, New = ifelse(
   (height>180) & (mass > 100) & (species =="Human"), "Big", "NotBig")
   )
```


## Gather 

takes multiple columns, and gathers them into key-value pairs 
- Make wide data longer
- Most important as analysis methods often prefer this form
- key = new name for values in columns
- value = new name for data values
- columns describe which columns to take

```{r}
# wide form
tempsData <- read_delim("data/cityTemps.txt", delim=" ")

tempsData %>% 
  gather(key=day, value=temp, 2:8)

newTempData<- tempsData %>% 
  gather(key=day, value=temp, sun:sat)

```


## Spread

takes two columns(key & value) and spreads in to multiple columns
- Make long data wider (opposite from gather)
- key = new column names
- value = value to spread out 

Also gives an easy method to seperate a column (or combine two columns) using `seperate()` and `unit()` 

```{r}
newTempData %>%
  spread(key=day, value=temp)

# Read a data
chicagoData <- read_csv("data/Chicago.csv") 

# default: remove=TRUE (remove the original variable)
chicagoData %>%
  separate(date, c("Day", "Month", "Year"), sep= "/")  

chicagoData %>%
  separate(date, c("Day", "Month", "Year"), sep= "/", convert = TRUE, remove = FALSE)

```


